Memory-Efficient Mixed-Precision Implementations\\for Robust Explicit Model Predictive Control

### 2019-07-29 emsoft-2019/camera_ready/Cont_Syn.tex
My pass on the final version.

The third input parameter is $\varepsilon_Q$. In~\autoref{sec:Error_Analysis} we
showed that if the user provides a precision $p$, then we can automatically
compute $\varepsilon_Q$.
Now, we want to \emph{derive} a suitable precision $p$, however, and face an
issue. In order to derive $p$ for some expression, Daisy and any other precision
tuning tool requires an error bound which should be satisfied. I.e. either a
precision $p$ is given, and we can compute an error bound, or the error bound is
given and we can derive a precision $p$. We cannot do both at the same time; the
problem would be underconstrained. We solve this chicken-and-egg issue by
requiring the user to provide $\varepsilon_Q$ as the input. 
%\Mahmoud{
%	\begin{remark}\label{rem:eps_Q}
%		Typical values for $\varepsilon_Q$ can range from $10^{-4}$ up-to $10^{-2}$. Increasing the $\varepsilon_Q$ leads into more memory save over $H$ and $K$, while requiring more precision for $F$ and $G$ (\autoref{tab:di}). Of course, too large values for $\varepsilon_Q$ can result in failure of RMPC design. However, this is usually not the case over reasonable choices for $\varepsilon_Q$.
%	\end{remark}}

We need to distribute the available `disturbance budget' among the two error
sources: error due to selecting the wrong region, which is given by the
quantization of $\hat{H}$ and $\hat{K}$, and the error due to
quantization of the control action, given by $\hat{F}$, $\hat{G}$.
The quantization of $\hat{H}$ and $\hat{K}$ is determined by
$\varepsilon_Q$, thus by providing this input value, the user effectively
chooses how much precision to allow for $\hat{H}$ and $\hat{K}$.
Note that this precision comes at the expense of the precision of
$\hat{F}$ and $\hat{G}$, i.e. the more accurately $\hat{H}$
and $\hat{K}$ are represented, the more approximate $\hat{F}$ and
$\hat{G}$ need to be and vice versa.
We note that one could straight-forwardly extend our algorithm to include an
outer loop which would explore different values for $\varepsilon_Q$. For our
experiments in~\autoref{sec:experiments}, we vary this value manually.

@@

Unlike the initial $\Delta$, $\varepsilon_Q$ needs to be strictly positive (line 4)
for mixed-precision tuning tools to work correctly. \Mahmoudd{Typical values for $\varepsilon_Q$ can range from $10^{-4}$ up to $10^{-2}$. Increasing $\varepsilon_Q$ leads into more memory save over $H$ and $K$ while requiring more precision for $F$ and $G$. Of course, too large values for $\varepsilon_Q$ can result in failure of the RMPC design. However, this is usually not the case over reasonable choices for $\varepsilon_Q$.}
Line 5 computes the width of the tubes $\varepsilon$ as shown in~\autoref{eq:epsilontot},
which will be used to bound the error due to selecting the wrong region
($\varepsilon_{A/D}$ is a constant parameter dependent on the specifics of the
converter).

@@

\smallskip
\noindent
\Mahmoudd{
\textbf{Termination of the algorithm.}
From a theoretical point of view, the error due to a wrong region selection does not necessarily converge since each new designed controller may differ from the previous ones. Therefore, it is not easy to give an upper bound over the number of times that the robust MPC design needs to be repeated. However, by rewriting \autoref{eq:RMPC_prob}, we have
\begin{align*}
\nu_0(x_0, \Delta),\ldots&,\nu_{N-1}(x_0,\Delta)=\\
argmin & \max_{w_0,\cdots,w_{N-1}} \sum_{i=0}^{N-1}(x_i^TQx_i+u_i^TRu_i) + x_N^TQ_Fx_N\nonumber\\
\quad  \text{s.t.} \quad & x_{i+1}=Ax_i+Bu_i + E w_i, \quad\forall i\in\{0,1,\cdots,N-1\}\nonumber\\
\quad \quad & u_i=\mu x_i+v_i, \quad\forall i\in\{0,1,\cdots,N-1\}\nonumber\\
\quad \quad & u_i\in\mathcal{U},x_i\in\mathcal{X},\quad \forall w_i\in\mathcal{W}_{\Delta},\,\,\forall i\in\{0,1,\cdots,N\}.
%&x_i\in\mathcal{X},\quad \forall w_i\in\mathcal{W},\quad\forall i\in\{0,1,\cdots,N\},
\end{align*}
Assuming the optimization has a unique solution for every $\Delta$, 
one may use the results on continuity of the $argmin$ function to compute an upper bound on the Lipschitz constant of the optimal input $\nu = \kappa(x)$ in Equation~\eqref{eq:affine_map} that holds for all $\Delta$. The computed bound will only depend on the dynamics of the system depicted in \autoref{eq:DSS} and hence can be used to evaluate the maximum error over incorrect region selection. This bound can be quite large, leading to a very conservative bound on the maximum number of iterations before convergence. From an experimental point of view, our observations show that using a reasonable $\varepsilon_{SAFE}$, convergence always reached after only a few iterations.}

\smallskip
\noindent
\Mahmoudd{
\textbf{Computational complexity.} It is shown in \cite{Alessio2009} that the number of regions $P$ in Equation~\eqref{eq:affine_map} is upper bounded by $2^q$, where $q$ is the total number of constraints in \autoref{eq:RMPC_prob}.
%In theory, one cannot find an upper bound for the maximum number of hyperplanes enclosing a region.
Assuming that each region has at most $\bar n_f$ hyperplanes, computation of the incorrect region selection error requires at most $i\times\bar n_f\times 2^q$ affine computations, where $i$ denotes the number of times the RMPC design updates $\Delta$ before convergence. In our experiments, the computation of the incorrect region selection error takes only a few minutes for our largest tested benchmarks. %\underline{\textbf{Rocco: What about complexity of Daisy's computations in terms of number of regions $P$ and $n$?}} 		
\\
The complexity of the precision tuning is linear with respect to the number of regions and hyperplanes. So, $(P+\bar n_f\times P)$ is an upper bound for the number of problems we need to encode in Daisy.
In each encoding, the uniform-precision detection takes linear number of steps before convergence. It starts from 1-bit and increases the precision one by one until the first valid configuration. The complexity of mixed-precision tuning with delta-debugging is $n\times log(n)$ in average, and bounded by $n^2$ in the worst case, with $n$ being the number of variables.}

>>>>

The third input parameter is $\varepsilon_Q$. In~\autoref{sec:Error_Analysis} we
showed that if the user provides a precision $p$, then we can automatically
compute $\varepsilon_Q$.
Now, we want to \emph{derive} a suitable precision $p$, however, and face an
issue. In order to derive $p$ for some expression, Daisy and any other precision
tuning tool requires an error bound which should be satisfied. That is, either a
precision $p$ is given, and we can compute an error bound, or the error bound is
given and we can derive a precision $p$. We cannot do both at the same time; the
problem would be underconstrained. We solve this chicken-and-egg problem by
requiring the user to provide $\varepsilon_Q$ as part of the input.
%\Mahmoud{
%	\begin{remark}\label{rem:eps_Q}
%		Typical values for $\varepsilon_Q$ can range from $10^{-4}$ up-to $10^{-2}$. Increasing the $\varepsilon_Q$ leads into more memory save over $H$ and $K$, while requiring more precision for $F$ and $G$ (\autoref{tab:di}). Of course, too large values for $\varepsilon_Q$ can result in failure of RMPC design. However, this is usually not the case over reasonable choices for $\varepsilon_Q$.
%	\end{remark}}

We need to distribute the available ``disturbance budget'' among the two error
sources: error due to selecting the wrong region, which is given by the
quantization of $\hat{H}$ and $\hat{K}$, and the error due to
quantization of the control action, given by $\hat{F}$, $\hat{G}$.
The quantization of $\hat{H}$ and $\hat{K}$ is determined by
$\varepsilon_Q$, thus by providing this input value, the user effectively
chooses how much precision to allow for $\hat{H}$ and $\hat{K}$.
Note that this precision comes at the expense of the precision of
$\hat{F}$ and $\hat{G}$, i.e. the more accurately $\hat{H}$
and $\hat{K}$ are represented, the more approximate $\hat{F}$ and
$\hat{G}$ need to be to fit into the disturbance budget, and vice versa.
We note that one could straight-forwardly extend our algorithm to include an
outer loop which would explore different values for $\varepsilon_Q$. For our
experiments in~\autoref{sec:experiments}, we vary this value manually.

@@

Unlike the initial $\Delta$, the parameter $\varepsilon_Q$ needs to be strictly positive (line 4)
for mixed-precision tuning tools to work correctly. 
Typical values for $\varepsilon_Q$ can range from $10^{-4}$ up to $10^{-2}$. 
\RM{repetition:}
Increasing $\varepsilon_Q$ leads to more memory savings over $H$ and $K$ while requiring more precision for $F$ and $G$. 
\RM{end}
Of course, a large value for $\varepsilon_Q$ can result in failure of the RMPC design. 
\RM{imprecise:}However, this is usually not the case over reasonable choices for $\varepsilon_Q$.
\RM{omit?}
Line 5 computes the width of the tubes $\varepsilon$ as shown in~\autoref{eq:epsilontot},
which will be used to bound the error due to selecting the wrong region
($\varepsilon_{A/D}$ is a constant parameter dependent on the specifics of the
converter).

@@

\smallskip
\noindent
\textbf{Termination of the algorithm.}
From a theoretical point of view, the error due to a wrong region selection does not necessarily converge since each new designed controller may differ from the previous ones. Therefore, it is not easy to give an upper bound over the number of times that the robust MPC design needs to be repeated. However, by rewriting \autoref{eq:RMPC_prob}, we have
\begin{align*}
\nu_0(x_0, \Delta),\ldots&,\nu_{N-1}(x_0,\Delta)=\\
\mathrm{argmin} & \max_{w_0,\cdots,w_{N-1}} \sum_{i=0}^{N-1}(x_i^TQx_i+u_i^TRu_i) + x_N^TQ_Fx_N\nonumber\\
\quad  \text{s.t.} \quad & x_{i+1}=Ax_i+Bu_i + E w_i, \quad\forall i\in\{0,1,\cdots,N-1\}\nonumber\\
\quad \quad & u_i=\mu x_i+v_i, \quad\forall i\in\{0,1,\cdots,N-1\}\nonumber\\
\quad \quad & u_i\in\mathcal{U},x_i\in\mathcal{X},\quad \forall w_i\in\mathcal{W}_{\Delta},\,\,\forall i\in\{0,1,\cdots,N\}.
%&x_i\in\mathcal{X},\quad \forall w_i\in\mathcal{W},\quad\forall i\in\{0,1,\cdots,N\},
\end{align*}
Assuming the optimization has a unique solution for every $\Delta$, 
one may use the results on continuity of the $\mathrm{argmin}$ function to compute an upper bound on the 
Lipschitz constant of the optimal input $\nu = \kappa(x)$ in Equation~\eqref{eq:affine_map} that holds for all $\Delta$. 
The computed bound will only depend on the dynamics of the system depicted in \autoref{eq:DSS} and hence can be used 
to evaluate the maximum error over incorrect region selection. 
This bound can be quite large, leading to a very conservative bound on the maximum number of iterations before convergence. 
From an experimental point of view, our observations show that for our range of parameters $\varepsilon_{\mathit{SAFE}}$, convergence was
always reached after only a few iterations.

\smallskip
\noindent
\textbf{Computational complexity.} 
It is shown in \cite{Alessio2009} that the number of regions $P$ 
in Equation~\eqref{eq:affine_map} is upper bounded by $2^q$, where $q$ is the total number of constraints in \autoref{eq:RMPC_prob}.
Assuming that each region has at most $N$ hyperplanes, computation of the incorrect region selection error requires at 
most $i\times N\times 2^q$ affine computations, where $i$ denotes the number of times the RMPC design updates $\Delta$ before convergence. 
In our experiments, the computation of the incorrect region selection error takes only a few minutes for our largest tested benchmarks. 
The complexity of the precision tuning is linear with respect to the number of regions and hyperplanes. 
So, $(P+ N\times P)$ is an upper bound for the number of problems we need to encode in Daisy.
In each encoding, the uniform-precision detection takes linear number of steps \RM{in what parameter?} before convergence. 
It starts from 1-bit and increases the precision one by one until the first valid configuration. 
The complexity of mixed-precision tuning with delta-debugging is $n\times log(n)$ in average, 
and bounded by $n^2$ in the worst case, where $n$ is the number of variables. \RM{how is $n$ related to the other parameters?}

### 2019-07-29 emsoft-2019/camera_ready/Error_Analysis.tex
class is updated into acmsmall
My pass on the final version.

We have to ensure that the sum of these two errors remain below the bound $\Delta$
used in the design of the robust explicit MPC:
\begin{align}
  \max\{\|v_{i}-v_{j}\|,\,&\forall i,j,\,\, i\neq j, \mathcal{R}_{i}\cap\mathcal{R}_{j}\neq\emptyset\} \nonumber\\
  &+  \max\{err(v_{i})_{p},\, i=1,2,\ldots,P\} \le \Delta.
  \label{eq:delta}
\end{align}
%where $i,j$ are the indexes of two distinct regions $\mathcal{R}_{i}$ and $\mathcal{R}_{j}$, 
The first term is the error of incorrect region selection. The maximum is taken over all neighboring regions 
$\mathcal R_i,\mathcal R_j$ (two regions are neighboring if their intersection is non-empty).
The function norm $\|v_{i}-v_{j}\|$ is taken by maximizing over all possible values of the state $x_k$. 
$err(v_{i})_{p}$ in
the second term captures the approximation error in the computation of the affine function $v_{i}$ when a fixed precision $p$ is used.
%$\mathit{neighbor}(\mathcal{R}_{i},\mathcal{R}_{j})$
%states that two regions share at least a point, $v_{i}$ is the control action
%for controller $i$, and $err(v_{i})_{p}$ captures the approximation error for
%computing the control action $v_{i}$ in fixed precision $p$.

@@

While our actual synthesis algorithm (\autoref{sec:Controller_Synthesis}) does
not compute the errors $\varepsilon_Q$ and $err(v_{i})_{p}$ explicitly, the
error verification is used internally by Daisy to determine a suitable precision
$p$ and can also be called explicitly without the precision optimization. \Mahmoudd{To encourage reproducible research, our implementation is made available online.\footnote{\url{https://github.com/rospoly/rmpc-daisy}}}
\Mahmoudd{\begin{remark}
	Note that we use MATLAB and Daisy solely for offline computations. The output of our algorithm contains (almost) all the information required for implementation of the verified controller over embedded systems. Therefore, the same savings can be expected when the controller is implemented on an actual embedded system.
\end{remark}}
\begin{comment}
\textcolor{red}{Sadegh: This is long comment on the content of this section. I cannot figure out how this line of reasoning is mapped to the implementation. We can discard my comment given that we are close to the deadline.}

>>>>

We have to ensure that the sum of these two errors remain below the bound $\Delta$
used in the design of the robust explicit MPC:
\begin{equation}
  \max\{\|v_{i}-v_{j}\|,\,\forall i,j,\,\, i\neq j, \mathcal{R}_{i}\cap\mathcal{R}_{j}\neq\emptyset\} +  \max\{\mathit{err}(v_{i})_{p},\, i=1,2,\ldots,P\} \le \Delta.
  \label{eq:delta}
\end{equation}
%where $i,j$ are the indexes of two distinct regions $\mathcal{R}_{i}$ and $\mathcal{R}_{j}$, 
The first term is the error of incorrect region selection. The maximum is taken over all neighboring regions 
$\mathcal R_i,\mathcal R_j$ (two regions are neighboring if their intersection is non-empty).
The function norm $\|v_{i}-v_{j}\|$ is taken by maximizing over all possible values of the state $x_k$ \RM{in $\mathcal{R}_i \cap \mathcal{R}_j$? Or all states?}. 
The term $\mathit{err}(v_{i})_{p}$ in
the second part captures the approximation error in the computation of the affine function $v_{i}$ when a fixed precision $p$ is used.
%$\mathit{neighbor}(\mathcal{R}_{i},\mathcal{R}_{j})$
%states that two regions share at least a point, $v_{i}$ is the control action
%for controller $i$, and $err(v_{i})_{p}$ captures the approximation error for
%computing the control action $v_{i}$ in fixed precision $p$.

@@

While our actual synthesis algorithm (\autoref{sec:Controller_Synthesis}) does
not compute the errors $\varepsilon_Q$ and $err(v_{i})_{p}$ explicitly, the
error verification is used internally by Daisy to determine a suitable precision
$p$ and can also be called explicitly without the precision optimization.

Note that we use MATLAB and Daisy solely for offline computations during the compilation of the controller.
At runtime, there is no additional cost.
% The output of our algorithm contains (almost) all the information required for implementation of the verified controller over embedded systems. 
% Therefore, the same savings can be expected when the controller is implemented on an actual embedded system.

\RM{what about the following commented-out comment of Sadegh?}
\begin{comment}
\textcolor{red}{Sadegh: This is long comment on the content of this section. I cannot figure out how this line of reasoning is mapped to the implementation. We can discard my comment given that we are close to the deadline.}

### 2019-07-29 emsoft-2019/camera_ready/Motiv.tex
My pass on the final version.

As a smaller example, consider the standard problem of designing a controller for an inverted pendulum depicted in~\autoref{fig:inverted_pendulum}~{(a)}.
The goal of the controller is to keep the pendulum at the vertical position while satisfying hard constraints on the state variables and control inputs.
A model of the system can be constructed using physical principles. 
After linearization and time discretization, the model is
	\begin{equation}
		\begin{bmatrix}
			 \theta_{k+1}\\
			\omega_{k+1}
		\end{bmatrix}=
		\begin{bmatrix}
			1 & T_s\\
			\frac{T_sg}{L}& (1-\frac{T_sb}{mL^2})		
		\end{bmatrix}
		\begin{bmatrix}
			\theta_k\\
			\omega_k
		\end{bmatrix}+
		\begin{bmatrix}
			0\\
			\frac{T_s}{mL^2}
		\end{bmatrix}u_k + w_k
		\label{eq:pendul_ss}
	\end{equation}
where $\theta_k$, $\omega_k$, and $u_k$ denote respectively the angular position, angular speed, 
and the input torque at time $kT_s$ with $T_s$ being an appropriate sampling time.
The disturbance $w_k\in\mathbb R^2$, bounded with a polyhedral set $w_k\in \mathcal W$, captures the modeling error due to 
linearization and discretization. 
The parameter $g=9.81 [m/s^2]$ is the gravitational acceleration, $m$ is the ball mass, $b$ is the rotational fraction coefficient, 
and $L$ is the length of the bar. 
Starting from an initial state $(\theta_0,\omega_0)$, the control goal is to converge to the equilibrium point 
$\theta=0, \omega=0$.
Additionally, we require the state constraints $\theta_k\in[-\pi,\pi]$ and $\omega_k\in[-\pi/8,\pi/8]$ to hold at all time instances.

\begin{figure}[t]
	\centering
  \begin{subfigure}[b]{0.33\columnwidth}
      \includegraphics[width=\textwidth]{Figs/inverted_pendulum.pdf}
      \vspace{0.05cm}
      \caption{}
  \end{subfigure}
  %~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
  \begin{subfigure}[b]{0.66\columnwidth}
      \includegraphics[width=\textwidth]{Figs/regs.jpg}
      \caption{}
  \end{subfigure}
	\caption{(a) Inverted pendulum; (b) 2D plot of polyhedral partitions for an EMPC applied to the inverted pendulum.}
	\label{fig:inverted_pendulum}
\end{figure}

@@

Figure \ref{fig:overview} gives a high-level overview of our proposed setup.
We start by selecting an initial bound $\Delta_0$ on the implementation error and enlarge $\mathcal W$ with $\Delta_0$ as $\mathcal W_{\Delta_0}$, where
\begin{equation}\label{eq:W_delta}
\mathcal W_{\Delta} := \{w_1+w_2 \,|\, w_1\in \mathcal W, \|w_2\|\le \Delta\}
\quad\text{ for all } \Delta\ge 0.
\end{equation}

For our example, we choose $\Delta_0=0.01$ and $\mathcal W = \{0\}$. 
We use Matlab to find a robust explicit MPC with disturbance set $\mathcal W_{\Delta_0}$. 
%
The output of robust EMPC given by Matlab decomposes the control domain into a
finite number of polyhedral domains together with an affine map for each domain. 
At each time step, the state vector $x_k$ is read (or estimated) and based on the domain
that it belongs to, the corresponding control output is computed. 
However, the polyhedral regions are implemented in finite precision so that
it is possible that due to quantization errors the domain is selected incorrectly, and therefore
a different affine map is computed for the control.
We refer to the resulting error as \emph{incorrect region selection} error.
% However, due to error by fixed-point implementation, it is possible that the domain is selected
% incorrectly. This causes an error term which we refer to it as incorrect region
% selection error.
In our example, we compute an incorrect region selection error $e = 0.019$, which is larger
than $\Delta_0=0.01$. We therefore increase the disturbance bound by a fixed amount $\varepsilon_{SAFE}=0.031$ to
$\Delta_1 = 0.05$ and enlarge the disturbance set as $\mathcal W_{\Delta_1}$.

%
Solving the robust explicit MPC problem for this enlarged disturbance set gives the new controller
with 14 regions, shown in~\autoref{fig:inverted_pendulum}(b).
A sample region with the corresponding affine map is shown below:
\begin{equation*}
%\resizebox{.9\hsize}{!}{$
\begin{bmatrix}
-0.005  &0.999\\
-0.998  &-0.049\\
0.005  &-0.999\\
0.998   &0.049
\end{bmatrix}x_k\leq 
\begin{bmatrix}
0.416\\
3.157\\
0.617\\
-0.0124
\end{bmatrix}
\Rightarrow
u(x_k)=\begin{bmatrix}
0.05&-9.67
\end{bmatrix}x_k+4.02%$}
\end{equation*}
where the output of robust EMPC at state $x_k$ is denoted by $u(x_k)$. 
The remaining mappings have a similar structure.
For this new controller, the incorrect region selection error is 
approximately $e = 0.019$, which is now below the disturbance bound $\Delta_1$,
so that we can continue with the precision assignment.

@@

Next, we use the precision tuning tool Daisy \cite{Daisy} to provide a mixed-precision 
implementation for all the parameters which satisfies the error bound $\Delta_1  - e = 0.05 - 0.019 = 0.031$.
The mixed-precision implementation returned by Daisy requires $6084$ bits.
Thus, for this introductory example, the mixed-precision implementation takes about $10\%$ less memory compared 
to the smallest uniform precision implementation that respects $\Delta_1$,
and about $57\%$ less memory than an implementation that uniformly uses
32 bits.

>>>>

As a simple example, consider the standard problem of designing a controller for an inverted pendulum depicted in~\autoref{fig:inverted_pendulum}~{(a)}.
The goal of the controller is to keep the pendulum at the vertical position while satisfying hard constraints on the state variables and control inputs.
A model of the system can be constructed using physical principles. 
After linearization and time discretization, the model is
	\begin{equation}
		\begin{bmatrix}
			 \theta_{k+1}\\
			\omega_{k+1}
		\end{bmatrix}=
		\begin{bmatrix}
			1 & T_s\\
			\frac{T_sg}{L}& (1-\frac{T_sb}{mL^2})		
		\end{bmatrix}
		\begin{bmatrix}
			\theta_k\\
			\omega_k
		\end{bmatrix}+
		\begin{bmatrix}
			0\\
			\frac{T_s}{mL^2}
		\end{bmatrix}u_k + w_k
		\label{eq:pendul_ss}
	\end{equation}
where $\theta_k$, $\omega_k$, and $u_k$ denote respectively the angular position, angular speed, 
and the input torque at time $kT_s$ with $T_s$ being an appropriate sampling time.
The disturbance $w_k\in\mathbb R^2$, bounded with a polyhedral set $w_k\in \mathcal W$, captures the modeling error due to 
linearization and discretization. 
The parameter $g=9.81 [m/s^2]$ is the gravitational acceleration, $m$ is the ball mass, $b$ is the rotational friction coefficient, 
and $L$ is the length of the bar. 
Starting from an initial state $(\theta_0,\omega_0)$, the control goal is to converge to the equilibrium point 
$\theta=0, \omega=0$.
Additionally, we require the state constraints $\theta_k\in[-\pi,\pi]$ and $\omega_k\in[-\pi/8,\pi/8]$ to hold at all time instances.

\begin{figure}[t]
	\centering
  \begin{subfigure}[b]{0.33\columnwidth}
      \includegraphics[width=\textwidth]{Figs/inverted_pendulum.pdf}
      \vspace{0.05cm}
      \caption{}
  \end{subfigure}
  %~ %add desired spacing between images, e. g. ~, \quad, \qquad, \hfill etc. 
  \begin{subfigure}[b]{0.66\columnwidth}
      \includegraphics[width=\textwidth]{Figs/regs.jpg}
      \caption{}
  \end{subfigure}
	\caption{(a) Inverted pendulum; (b) 2D plot of polyhedral partitions for an EMPC for the inverted pendulum.}
	\label{fig:inverted_pendulum}
\end{figure}

@@

Figure \ref{fig:overview} gives a high-level overview of our proposed setup.
We start by selecting an initial bound $\Delta_0$ on the implementation error and enlarge $\mathcal W$ with $\Delta_0$ as $\mathcal W_{\Delta_0}$, where
we define
\begin{equation}\label{eq:W_delta}
\mathcal W_{\Delta} := \{w_1+w_2 \,|\, w_1\in \mathcal W, \|w_2\|\le \Delta\}
\quad\text{ for all } \Delta\ge 0.
\end{equation}

For our example, we choose $\Delta_0=0.01$ and $\mathcal W = \{0\}$. 
We use Matlab to find a robust explicit MPC with disturbance set $\mathcal W_{\Delta_0}$. 
%
The output of robust EMPC given by Matlab decomposes the control domain into a
finite number of polyhedral domains together with an affine map for each domain. 
At each time step, the state vector $(\theta_k, \omega_k)$ is read (or estimated), and based on the polyhedral domain
that it belongs to, the corresponding affine map is computed as the control output.

Since the polyhedral regions are implemented in finite precision, 
it is possible that due to quantization errors, the polyhedral domain is selected incorrectly, and therefore
a different affine map is computed for the control.
We refer to the resulting error as \emph{incorrect region selection} error.
In our example, we compute an incorrect region selection error $e = 0.019$, which is larger
than $\Delta_0=0.01$. We therefore increase the disturbance bound by a fixed amount $\varepsilon_{SAFE}=0.031$ to
$\Delta_1 = 0.05$ and enlarge the disturbance set as $\mathcal W_{\Delta_1}$.

%
Solving the robust explicit MPC problem for this enlarged disturbance set gives the new controller
with 14 regions, shown in~\autoref{fig:inverted_pendulum}(b).
A sample region with the corresponding affine map is shown below:
\begin{equation*}
%\resizebox{.9\hsize}{!}{$
\begin{bmatrix}
-0.005  &0.999\\
-0.998  &-0.049\\
0.005  &-0.999\\
0.998   &0.049
\end{bmatrix}
\begin{bmatrix}
\theta_k\\
\omega_k
\end{bmatrix}
%
\leq 
\begin{bmatrix}
0.416\\
3.157\\
0.617\\
-0.0124
\end{bmatrix}
\Rightarrow
u(\begin{bmatrix} \theta_k\\ \omega_k \end{bmatrix})=\begin{bmatrix}
0.05&-9.67
\end{bmatrix}\begin{bmatrix}\theta_k\\ \omega_k\end{bmatrix} +4.02%$}
\end{equation*}
where the output of robust EMPC at state $(\theta_k, \omega_k)$ is denoted by $u(\theta_k, \omega_k)$. 
The remaining mappings have a similar structure.
For this new controller, the incorrect region selection error is 
approximately $e = 0.019$, which is now below the disturbance bound $\Delta_1$,
so that we can continue with the precision assignment.

@@

We can implement the controller using uniform $32$ bit integers, or by selecting a uniform bitwidth.
These take, respectively, $XX$ and $YY$ bits.
\RM{how many bits do we need for the uniform prec impl? how many bits for 32 bits?} 
But we can do better.
We use the precision tuning tool Daisy \cite{Daisy} to provide a mixed-precision 
implementation for all the parameters which satisfies the error bound $\Delta_1  - e = 0.05 - 0.019 = 0.031$.
The mixed-precision implementation returned by Daisy requires $6084$ bits.
Thus, for this example, the mixed-precision implementation takes about $10\%$ less memory compared 
to the smallest uniform precision implementation that respects $\Delta_1$,
and about $57\%$ less memory than an implementation that uniformly uses 32 bits.

### 2019-07-29 emsoft-2019/camera_ready/RMPC_Daisy.tex
My pass on the final version.

The implementation of the controller stores matrices $F_i,G_i,H_i,K_i$ for all $i\in\{1,2,\ldots,P\}$. 
At each time step $k$, the current state $x_k$ is used to detect the polyhedral region 
such that $H_i x_k\le K_i$. Then the control action $v_k = F_i x_k + G_i$ is computed and applied to the system.
Several algorithms are proposed in the literature \cite{Mnnigmann:2011,Jones:2006} to find the right polyhedral region $i$ to which the state $x_k$ belong. 
The most straightforward technique is a linear search over all polyhedral regions (Fig.~\ref{lst:caseof}).
\Mahmoudd{
\begin{remark}
	The total time to evaluate the PWA function in \autoref{eq:affine_map} for a given $x_k$ consists of two steps: (a) selecting the region $\mathcal R_i$ such that $x_k\in \mathcal R_i$ and (b) evaluating the corresponding affine map $F_i x_k +G_i$.
	Evaluating the affine map is straightforward and requires less computational effort compared to the region selection step. Direct implementation of region selection (\autoref{lst:caseof}) is simple, but inefficient. More efficiency can be achieved by leveraging binary search tree structures (e.g. \cite{Mnnigmann:2011}).\\
	In general, explicit MPC is well-known for its high-speed implementation over embedded systems with low computational power. The authors in \cite{Johansen:2007} show that explicit MPCs with hundreds of regions can be implemented on application specific integrated circuits (ASIC) with about $20000$ gates, leading to computation times in the order of $1\mu s$. It is shown in \cite{Bemporad:2006} that for typical problems evaluating the explicit MPC takes significantly shorter time in comparison to solving on-line quadratic programs.
\end{remark}}

\Mahmoudd{ 
 \begin{remark}
 It should be noticed that EMPC is a class of controllers specifically proposed for linear time invariant (LTI) systems. To cover nonlinear systems, one approach is to linearize them around an appropriately selected equilibrium point of the system and to model the differences as a bounded disturbance as in \autoref{eq:RMPC_prob}. Taking the upper bound over nonlinearities into account will result in enlarging $\Delta$ which characterizes $\mathcal{W}_{\Delta}$. For reasonable values of $\Delta$, in principle RMPC design should work successfully. Further, it should be noticed that the rest of analysis in the sequel (over MATLAB and Daisy) do not rely on the LTI property of the system under study and could be adjusted for other families of controllers.
 %However, such analysis are beyond the scope of this paper as (perturbed) LTI systems are considered as an important class of dynamical systems.
\end{remark}}

@@

To keep the implementation cost down, low-end microcontrollers usually have
limited memory and computational power. Therefore, the implementation of the above
explicit MPC on microcontrollers will bring some issues since all regions'
matrices ($H_i$s and $K_i$s) and activation functions ($F_i$s and $G_i$s) have
to be stored on the microcontroller. Our paper presents an
approach for minimizing the memory usage of the implementation while satisfying
the hard constraints on the states.
%\section{Finite Precision Implementation}
%\Sadegh{To be included by Rocco}

>>>>

The implementation of the controller stores matrices $F_i,G_i,H_i,K_i$ for all $i\in\{1,2,\ldots,P\}$. 
At each time step $k$, the current state $x_k$ is used to detect the polyhedral region 
such that $H_i x_k\le K_i$. Then the control action $v_k = F_i x_k + G_i$ is computed and applied to the system.
Several algorithms are proposed in the literature \cite{Mnnigmann:2011,Jones:2006} to find the right polyhedral region $i$ to which the state $x_k$ belong. 
The most straightforward technique is a linear search over all polyhedral regions (Fig.~\ref{lst:caseof}).
More efficiency can be achieved by using binary search tree structures (e.g., \cite{Mnnigmann:2011}).

% The total time to evaluate the PWA function in \autoref{eq:affine_map} for a given $x_k$ consists of two steps: 
% (a) selecting the region $\mathcal R_i$ such that $x_k\in \mathcal R_i$ and 
% (b) evaluating the corresponding affine map $F_i x_k +G_i$.
% Evaluating the affine map is straightforward and requires less computational effort compared to the region selection step. 
% Direct implementation of region selection (\autoref{lst:caseof}) is simple, but inefficient.

To keep the implementation cost down, low-end microcontrollers usually have limited computational power and memory. 
In general, explicit MPC is well-known for its high-speed implementation in embedded systems with low computational power. 
Johansen et al.~\cite{Johansen:2007} show that explicit MPCs with hundreds of regions can be implemented on application specific integrated circuits (ASIC) 
with about $20000$ gates, leading to computation times in the order of $1\mu s$. 
It is shown in \cite{Bemporad:2006} that for typical problems evaluating the explicit MPC takes significantly less
time in comparison to solving on-line quadratic programs.
In this paper, we present an approach for minimizing the memory usage of the implementation while satisfying
the hard constraints on the states, thus ensuring the controllers can be implemented in low-memory microcontrollers.

While we focus on linear time invariant systems, EMPC is also applicable to nonlinear systems by linearization
around an appropriately selected equilibrium point and modeling the error in the linearization as a disturbance.
% 
% It should be noticed that EMPC is a class of controllers specifically proposed for linear time invariant (LTI) systems. 
% To cover nonlinear systems, one approach is to linearize them around an appropriately selected equilibrium point of the system and 
% to model the differences as a bounded disturbance as in \autoref{eq:RMPC_prob}. 
% This results in enlarging $\Delta$ which characterizes $\mathcal{W}_{\Delta}$. 
% For reasonable values of $\Delta$, in principle RMPC design should work successfully. 
Note that the rest of our analysis in the sequel (over MATLAB and Daisy) do not 
rely on the LTI property of the system under study and could be adjusted for other families of controllers
as long as they are presented as piecewise affine functions.

@@

### 2019-07-29 emsoft-2019/camera_ready/conclusion.tex
My pass on the final version.

We have described an automatic technique to tune an explicit MPC
controller that implements the controller in mixed-precision fixed-point arithmetic
while ensuring that the resulting loss of precision does not invalidate the
constraints of the original control problem.
Our technique models potential fixed-point imprecision explicitly as a disturbance
term, and uses robust explicit MPC to design a controller.
A static error analysis and a mixed-precision tuning tool is then 
used to find an efficient implementation of the controller function.
The implementation maintains the numerical error to within
the disturbance bounds used by the robust controller.
In experiments, we show that our technique can yield significant savings in memory,
ranging up to 20\%. 
In addition, the static analysis scales to controllers with hundreds of regions.\\
One exciting direction for future work would be to investigate the effect of stochastic disturbances on the proposed methodology and find a scheme to reduce statistically the memory footprint of the controller implementation.

>>>>

We have described an automatic technique to tune an explicit MPC
controller.
Our technique implements the controller in mixed-precision fixed-point arithmetic
while ensuring that the resulting loss of precision does not invalidate the
constraints of the original control problem.
We model potential fixed-point imprecision explicitly as a disturbance
term, and uses robust explicit MPC to design a controller.
A static error analysis and a mixed-precision tuning tool is then 
used to find an efficient implementation of the controller function.
The implementation maintains the numerical error to within
the disturbance bounds used by the robust controller.
In experiments, we show that our technique can yield significant savings in memory,
ranging up to 20\% in our experiments. 
In addition, the static analysis scales to controllers with hundreds of regions.
One direction for future work would be to investigate the effect of stochastic disturbances on the proposed methodology and find a 
scheme to statistically reduce the memory footprint of the controller implementation.
